---
title: "03_demultiplex"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = './')

library(rmarkdown)
library(tidyverse)
library(Seurat)
library(Signac)
library(ggplot2)
library(patchwork)
set.seed(1234)

### 

source(file =file.path("00_parameters.R"))
source(file = paste0(projPath, "/00_Rfunctions.R"))
# source(file =file.path("preprocessing.R"))

```

## Filter doublets and non-cells
### Construct metadata inclduing barcode stats and set thresholds (low) 
For fragments passing filter and TSS fragments. Will be further qualified by hashtags below.

```{r, fig.height=12, fig.width=12}
########## Filter the barcodes
cat("*** Reading barcode statistics files \n")

metadata_file      <- paste0(outsPath, '/singlecell.csv')
 if (!file.exists(metadata_file)) {stop(paste0("Metadata file does not exist: ",metadata_file))}

metadata = read.csv(metadata_file, header = 1)
metadata = metadata[2:nrow(metadata),]

metadata$logUMI = log10(metadata$passed_filters + 1)
metadata$promoter_ratio = (metadata$promoter_region_fragments+1) / (metadata$passed_filters + 1)
metadata$peak_region_ratio = (metadata$peak_region_fragments+1) / (metadata$passed_filters + 1)
  metadata$enhancer_region_ratio = (metadata$enhancer_region_fragments+1) / (metadata$passed_filters + 1)
metadata$EnhPromDNAseHyp_ratio = (metadata$on_target_fragments+1) / (metadata$passed_filters + 1)
metadata$FRiP<-metadata$peak_region_ratio*100

metadata$is_cellranger_cell_barcode<-metadata$is__cell_barcode

#Subset by thresholds:
metadata<- metadata %>%
  mutate(., is__cell_barcode = ifelse(
    #    passed_filters > cutoff_reads_min&
   # passed_filters > 10^cutoff_reads_min &  
  #    enhancer_region_ratio > cutoff_enhancer_ratio &
      passed_filters < 10^cutoff_reads_max &
  on_target_fragments > cutoff_fragmentsInRegions,
      "1", "0"))
table("MyFilter"=metadata$is__cell_barcode, "cellranger"=metadata$is_cellranger_cell_barcode)
```

##
```{r}
#############

p1 <- ggplot(data = metadata,aes(x=logUMI,y=EnhPromDNAseHyp_ratio,col=factor(is_cellranger_cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is_cellranger_cell_barcode)))),"FALSE")) +
  theme(legend.title = element_blank())+
  theme(legend.position="bottom",text=element_text(size=20)) +ggtitle("Cellranger cells")+ylim(c(0,1))



p2<-ggplot(data = metadata,aes(x=logUMI,y=EnhPromDNAseHyp_ratio,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is__cell_barcode)))),
                              "FALSE")) +
  theme(legend.position="bottom",text=element_text(size=20)) +
   geom_hline(yintercept = cutoff_enhancer_ratio,  linetype="dashed",col="red")+
   geom_vline(xintercept = (cutoff_reads_min),  linetype="dashed",col="red" )+
  geom_vline(xintercept = (cutoff_reads_max),  linetype="dashed",col="red" )+
  ggtitle("Cell calling by Enhancer Region Fragments")+
  theme(legend.title = element_blank())+ylim(c(0,1))

p1+p2 & theme(legend.position = 'bottom')
#plot_layout(guides = "collect")

####FrIP:
p_fripCellranger<-ggplot(data = metadata,aes(x=log10(passed_filters),y=FRiP,col=factor(is_cellranger_cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is_cellranger_cell_barcode)))),
                              "FALSE")) +
  ggtitle("Filtered by passed_filters fragments")+
  theme(legend.title = element_blank())

p_frip<-ggplot(data = metadata,aes(x=log10(passed_filters),y=FRiP,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is__cell_barcode)))),
                              "FALSE")) +
  ggtitle("Filtered by passed_filters fragments")+
  theme(legend.title = element_blank())

p_fripCellranger+p_frip& theme(legend.position = 'bottom')
#plot_layout(guides = "collect")


###
ggplot(data = metadata,aes(x=logUMI,y=(on_target_fragments),col=factor(is__cell_barcode))) + 
  geom_point(size=0.1)+ ylim(0,10*cutoff_fragmentsInRegions)


ggplot(data = metadata,aes(x=logUMI,y=on_target_fragments,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1)+ ylim(0,50*cutoff_fragmentsInRegions)
```



# Filter out Doublets and negatives using hashtags:
### Hash-counts for filtered barcodes
```{r, eval=T, fig.width=12, fig.height=20}
mtx_path <- paste0(hashPath, "/featurecounts.mtx" )
feature_path <- paste0(hashPath, "/featurecounts.genes.txt" )
barcode_path <- paste0(hashPath, "/featurecounts.barcodes.txt")
 
features <- readr::read_tsv(feature_path, col_names = F) %>% tidyr::unite(hto) 
barcodes <- readr::read_tsv(barcode_path, col_names = F) %>% tidyr::unite(barcode) 

hto_t <- Matrix::readMM(mtx_path) %>%
  magrittr::set_colnames(features$hto) %>%
  magrittr::set_rownames(paste0(barcodes$barcode, "-1"))

hto_unf<-t(as.matrix((hto_t)))

#@Oliver: instead of "metadata" use the confined list of cell barcodes you have
hto <- hto_unf[, intersect(metadata %>%
                            filter(is__cell_barcode==1) %>%
                             pull(barcode), colnames(hto_unf))]
#hto<-hto_unf
hto<-hto[,colSums(hto)<10000]
hto<-hto[-8,] # since very few sorted to scCT, very few cells found by demux -  it distorts the HTO quant. better to remove it all together.
hto<-hto[-3,] # since none sorted for scCUT&Tag # see labguru experiment.
dim(hto)



```
```{r}
metadata<- left_join(metadata, 
                     colSums(hto_unf) %>%
                       enframe(name="barcode", value="colSums_hto"))

metadata %>%
  filter(!is.na(colSums_hto))%>%
  ggplot(., aes(passed_filters, colSums_hto, col=factor(is_cellranger_cell_barcode)))+geom_point(size=0.4)+
  theme(legend.title = element_blank())+
  scale_x_log10()+scale_y_log10() +ggtitle(paste0(ExptName))
```


#### Create seruat HTO object
```{r}
#Create seurat object:
meta.data<-metadata
rownames(meta.data)<-NULL
meta.data<-meta.data %>% column_to_rownames("barcode")


seur_hto<-CreateSeuratObject(counts=hto, project = "hto" ,assay="HTO", min.features = 1,  min.cells = 1,
                             meta.data = meta.data)




seur_hto<-Seurat::NormalizeData(seur_hto, normalization.method = "CLR", assay="HTO")
seur_hto
```


### Run Demux
```{r}
  seur_hto<-MULTIseqDemux(seur_hto ,maxiter = 100,autoThresh = T, qrange =seq(from = 0.1, to = 0.99999, by = 0.05) )

# # Set quantile from above plot.
 # seur_hto<-HTODemux(seur_hto, positive.quantile = .99999999)
 # table(seur_hto$HTO_classification.global)
 # 

table(seur_hto$MULTI_ID, seur_hto$is_cellranger_cell_barcode)
seur_hto$primary_ID<-seur_hto$MULTI_ID
###############################
# Second Pass MULTI-seq DEMUX:#
###############################

secondPass<-MULTIseqDemux(
  subset(seur_hto, subset = MULTI_ID %in% c( "Negative")))

table(secondPass$MULTI_ID, secondPass$is_cellranger_cell_barcode)


thirdPass<-MULTIseqDemux(
  subset(secondPass, subset = MULTI_ID %in% c( "Negative")), autoThresh = T)

table(thirdPass$MULTI_ID, thirdPass$is_cellranger_cell_barcode)


seur_hto<-
merge(subset(seur_hto, subset = MULTI_ID!="Negative"), 
      subset(secondPass, subset = MULTI_ID!="Negative")
      )

seur_hto<-merge(subset(seur_hto, subset = MULTI_ID!="Negative"), 
      thirdPass
      )

table(seur_hto$MULTI_ID, seur_hto$is_cellranger_cell_barcode)
seur_hto

```

### plot MULTIseqDemux
```{r, fig.height=12, fig.width=12}
ggplot(seur_hto@meta.data, aes(x=nCount_HTO, y=EnhPromDNAseHyp_ratio, col=factor(MULTI_ID)))+
  geom_point()+
   scale_x_log10()+scale_y_log10()

ggplot(seur_hto@meta.data, aes(x=passed_filters, y=nCount_HTO, col=factor(MULTI_classification)))+
  geom_point()+
   scale_x_log10()+scale_y_log10()

RidgePlot(seur_hto, features=rownames(seur_hto[1:4]), ncol = 2, group.by="MULTI_classification")






```
### plot MULTIseqDemux
```{r, fig.height=12, fig.width=12}
RidgePlot(seur_hto, features=rownames(seur_hto[5:8]), ncol = 2, group.by="MULTI_ID")
```

### plot nCount_HTO
```{r}
ggplot(seur_hto@meta.data, aes(x=nCount_HTO, y=EnhPromDNAseHyp_ratio, col=factor(MULTI_ID)))+
  geom_point()+
   scale_x_log10()+scale_y_log10()

ggplot(seur_hto@meta.data, aes(x=passed_filters, y=nCount_HTO, col=factor(MULTI_ID)))+
  geom_point()+
   scale_x_log10()+scale_y_log10()
```

### tSNE and plotting of HTOs
```{r, fig.width=7, fig.height=7}

Idents(seur_hto) <- "MULTI_ID"
hto.dist.mtx <- as.matrix(dist(t(GetAssayData(object = seur_hto, assay = "HTO"))))
seur_hto <- RunTSNE(seur_hto, distance.matrix = hto.dist.mtx)


DimPlot(seur_hto,group.by = "MULTI_ID",pt.size = 1,cols= c("polychrome"))+ggtitle("MULTI_ID")#+NoLegend()


vln1<-VlnPlot(seur_hto, features=c("nCount_HTO", "passed_filters"), log = T)
vln2<-VlnPlot(seur_hto, features=c("EnhPromDNAseHyp_ratio", "FRiP"), log = F)
vln1/vln2

FeaturePlot(seur_hto, feature=c("nCount_HTO", "passed_filters","HTO_margin", "nCount_HTO","EnhPromDNAseHyp_ratio", "FRiP"))
```

### add Singlet info to metadata and plot
```{r, fig.width=12, fig.height=4, eval=F}



metadataFINAL<-
seur_hto@meta.data %>% rownames_to_column("barcode") %>%
  mutate(.,singlet = ifelse(
    MULTI_ID %in% c("Negative", "Doublet", "0"),
    "0", "1")) %>% 
  right_join(.,metadata) %>%
  mutate(is__cell_barcode=replace_na(singlet,0))

table("Singlet"=metadataFINAL$is__cell_barcode, "cellranger"=metadataFINAL$is_cellranger_cell_barcode)

### Plot:
###########
p3<-ggplot(data = metadataFINAL,aes(x=logUMI,y=EnhPromDNAseHyp_ratio,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadataFINAL$is__cell_barcode)))),
                              "FALSE")) +
  theme(legend.position="bottom",text=element_text(size=20)) +
   geom_hline(yintercept = cutoff_enhancer_ratio,  linetype="dashed",col="red")+
   geom_vline(xintercept = cutoff_reads_min,  linetype="dashed",col="red" )+
     geom_vline(xintercept = cutoff_reads_max,  linetype="dashed",col="red" )+
  ggtitle("Singlets from HTO")+
  theme(legend.title = element_blank())+ylim(c(0,1))


p1+p2+p3 & theme(legend.position = 'bottom')
#+ plot_layout(guides = "collect")


```

## Make signac object
```{r}

saveRDS(metadataFINAL, paste0(projPath,"/", ExptName, "_metadataObject.RDS"))
#saveRDS(metadata, paste0(projPath,"/", ExptName, "_metadataObject.RDS"))

```



## Make new bam-file for with only called cells
### R script to print filtered_cell_barcodes
```{r sinto_tsv}
dir.create(paste0(projPath,"/bam"))

cells<-
  # seur_hto@meta.data %>% rownames_to_column("barcode") %>%
  #  mutate(., is__cell_barcode = ifelse(
  #    HTO_classification.global =="Singlet",
  #    "1", "0")) %>%
 metadataFINAL %>%
  filter(is__cell_barcode=="1")%>% 
  dplyr::select(barcode, is__cell_barcode)

cells$is__cell_barcode<-paste0(ExptName, "_FilteredCells")

write_tsv(x = cells, file=paste0(projPath,"/bam/FilteredCellBarcodes.tsv"), col_names = FALSE)
```



## go to chapter 04.


```{r}
sessionInfo()

```
