---
title: "03_demultiplex"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = './')

library(rmarkdown)
library(tidyverse)
library(Seurat)
library(Signac)
library(ggplot2)
library(patchwork)
set.seed(1234)

### 

source(file =file.path("00_parameters.R"))
source(file = paste0(projPath, "/00_Rfunctions.R"))
```

## Filter doublets and non-cells
### Construct metadata inclduing barcode stats and set thresholds (low) 
For fragments passing filter and TSS fragments. Will be further qualified by hashtags below.

```{r, fig.height=12, fig.width=12}
########## Filter the barcodes
cat("*** Reading barcode statistics files \n")

metadata_file      <- paste0(outsPath, '/singlecell.csv')
 if (!file.exists(metadata_file)) {stop(paste0("Metadata file does not exist: ",metadata_file))}

metadata = read.csv(metadata_file, header = 1)
metadata = metadata[2:nrow(metadata),]
metadata$logUMI = log10(metadata$passed_filters + 1)
metadata$promoter_ratio = (metadata$promoter_region_fragments+1) / (metadata$passed_filters + 1)
metadata$peak_region_ratio = (metadata$peak_region_fragments+1) / (metadata$passed_filters + 1)
metadata$enhancer_region_ratio = (metadata$enhancer_region_fragments+1) / (metadata$passed_filters + 1)
metadata$FRiP<-metadata$peak_region_ratio*100

metadata$is_cellranger_cell_barcode<-metadata$is__cell_barcode

#Subset by thresholds:
metadata<- metadata %>%
  mutate(., is__cell_barcode = ifelse(
  #      passed_filters > cutoff_reads_min,
   passed_filters > 10^cutoff_reads_min &  
     enhancer_region_ratio > cutoff_enhancer_ratio &
     passed_filters < 10^cutoff_reads_max,
      "1", "0"))
table("MyFilter"=metadata$is__cell_barcode, "cellranger"=metadata$is_cellranger_cell_barcode)
```

##
```{r}
#############

p1 <- ggplot(data = metadata,aes(x=logUMI,y=enhancer_region_ratio,col=factor(is_cellranger_cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is_cellranger_cell_barcode)))),"FALSE")) +
  theme(legend.title = element_blank())+
  theme(legend.position="bottom",text=element_text(size=20)) +ggtitle("Cellranger cells")+ylim(c(0,0.5))



p2<-ggplot(data = metadata,aes(x=logUMI,y=enhancer_region_ratio,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is__cell_barcode)))),
                              "FALSE")) +
  theme(legend.position="bottom",text=element_text(size=20)) +
   geom_hline(yintercept = cutoff_enhancer_ratio,  linetype="dashed",col="red")+
   geom_vline(xintercept = (cutoff_reads_min),  linetype="dashed",col="red" )+
  geom_vline(xintercept = (cutoff_reads_max),  linetype="dashed",col="red" )+
  ggtitle("Cell calling by Enhancer Region Fragments")+
  theme(legend.title = element_blank())+ylim(c(0,0.5))

p1+p2 & theme(legend.position = 'bottom')
#plot_layout(guides = "collect")

####FrIP:
p_fripCellranger<-ggplot(data = metadata,aes(x=log10(passed_filters),y=FRiP,col=factor(is_cellranger_cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is_cellranger_cell_barcode)))),
                              "FALSE")) +
  ggtitle("Filtered by passed_filters fragments")+
  theme(legend.title = element_blank())

p_frip<-ggplot(data = metadata,aes(x=log10(passed_filters),y=FRiP,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is__cell_barcode)))),
                              "FALSE")) +
  ggtitle("Filtered by passed_filters fragments")+
  theme(legend.title = element_blank())

p_fripCellranger+p_frip& theme(legend.position = 'bottom')
#plot_layout(guides = "collect")


###

```



# Filter out Doublets and negatives using hashtags:
### Hash-counts for filtered barcodes
```{r, eval=T, fig.width=12, fig.height=20}
mtx_path <- paste0(hashPath, "/featurecounts.mtx" )
feature_path <- paste0(hashPath, "/featurecounts.genes.txt" )
barcode_path <- paste0(hashPath, "/featurecounts.barcodes.txt")
 
features <- readr::read_tsv(feature_path, col_names = F) %>% tidyr::unite(hto) 
barcodes <- readr::read_tsv(barcode_path, col_names = F) %>% tidyr::unite(barcode) 

hto_t <- Matrix::readMM(mtx_path) %>%
  magrittr::set_colnames(features$hto) %>%
  magrittr::set_rownames(paste0(barcodes$barcode, "-1"))

hto_unf<-t(as.matrix((hto_t)))

#@Oliver: instead of "metadata" use the confined list of cell barcodes you have
hto <- hto_unf[, intersect(metadata %>%
                            filter(is__cell_barcode==1) %>%
                             pull(barcode), colnames(hto_unf))]

dim(hto)

```

```{r}
metadata<- left_join(metadata, 
                     colSums(hto_unf) %>%
                       enframe(name="barcode", value="colSums_hto"))

metadata %>%
  filter(!is.na(colSums_hto))%>%
  ggplot(., aes(passed_filters, colSums_hto, col=factor(is_cellranger_cell_barcode)))+geom_point(size=0.4)+
  theme(legend.title = element_blank())+
  scale_x_log10()+scale_y_log10() +ggtitle(paste0(ExptName))
```

#### Create seruat HTO object
```{r}
#Create seurat object:
seur_hto<-CreateSeuratObject(counts=hto, project = "hto" ,assay="HTO", min.features = 1,  min.cells = 1)


seur_hto<-NormalizeData(seur_hto, normalization.method = "CLR", assay="HTO")
seur_hto
```

### Plot doublet and negative proportion to find fitting quantile
```{r}
### Plot doublet and negative proportion to find fitting quantile
dat <- map(seq(.999,1,length.out =10), function(x) HTODemux(seur_hto, positive.quantile = x))
res <- map_dfr(dat, function(x) table(x$hash.ID)) %>% mutate(quant=seq(.999,1,length.out =10))
res %>% pivot_longer(-quant) %>% 
  filter(quant!=1) %>% 
  ggplot() + aes(x=quant, y=value, color=name) + geom_point() + geom_line()
```

### Run Demux
```{r}

# Set quantile from above plot.
seur_hto<-HTODemux(seur_hto, positive.quantile = 0.999)
table(seur_hto$HTO_classification.global)

seur_hto<-AddMetaData(seur_hto, metadata %>% column_to_rownames("barcode"))
```

### plot nCount_HTO
```{r}
ggplot(seur_hto@meta.data, aes(x=nCount_HTO, y=enhancer_region_ratio, col=factor(hash.ID)))+
  geom_point()+
   scale_x_log10()+scale_y_log10()

ggplot(seur_hto@meta.data, aes(x=passed_filters, y=nCount_HTO, col=factor(hash.ID)))+
  geom_point()+
   scale_x_log10()+scale_y_log10()
```

### tSNE and plotting of HTOs
```{r, fig.width=7, fig.height=7}

Idents(seur_hto) <- "hash.ID"
hto.dist.mtx <- as.matrix(dist(t(GetAssayData(object = seur_hto, assay = "HTO"))))
seur_hto <- RunTSNE(seur_hto, distance.matrix = hto.dist.mtx)


DimPlot(seur_hto,group.by = "hash.ID",pt.size = 1,cols= c("polychrome"))+ggtitle("hash.ID")#+NoLegend()


DimPlot(seur_hto, group.by="HTO_maxID")+ggtitle("HTO_maxID")

vln1<-VlnPlot(seur_hto, features=c("nCount_HTO", "passed_filters"), log = T)
vln2<-VlnPlot(seur_hto, features=c("enhancer_region_ratio", "FRiP"), log = F)
vln1/vln2

FeaturePlot(seur_hto, feature=c("nCount_HTO", "passed_filters","HTO_margin", "nCount_HTO","enhancer_region_ratio", "FRiP"))
```

### add Singlet info to metadata and plot
```{r, fig.width=12, fig.height=4}
metadata<-
seur_hto@meta.data %>% rownames_to_column("barcode") %>%
  right_join(.,metadata) %>%
  mutate(.,singlet = ifelse(
    HTO_classification.global =="Singlet",
    "1", "0")) %>% 
  mutate(is__cell_barcode=replace_na(singlet,0))

table("Singlet"=metadata$is__cell_barcode, "cellranger"=metadata$is_cellranger_cell_barcode)

### Plot:
###########
p3<-ggplot(data = metadata,aes(x=logUMI,y=enhancer_region_ratio,col=factor(is__cell_barcode))) + 
  geom_point(size=0.1) + 
  scale_color_manual(values=c("black","gold"),
                     labels=c(paste("TRUE",sum(as.numeric(as.character(metadata$is__cell_barcode)))),
                              "FALSE")) +
  theme(legend.position="bottom",text=element_text(size=20)) +
   geom_hline(yintercept = cutoff_enhancer_ratio,  linetype="dashed",col="red")+
   geom_vline(xintercept = log10(cutoff_reads_min),  linetype="dashed",col="red" )+
  ggtitle("Singlets from HTO")+
  theme(legend.title = element_blank())+ylim(c(0,0.5))


p1+p2+p3 & theme(legend.position = 'bottom')
#+ plot_layout(guides = "collect")


```

## Make signac object
```{r}
meta.data<-seur_hto@meta.data %>% rownames_to_column("barcode") %>%
  mutate(., is__cell_barcode = ifelse(
    HTO_classification.global =="Singlet",
      "1", "0"))
saveRDS(meta.data, paste0(projPath,"/", ExptName, "_metadataObject.RDS"))
#saveRDS(metadata, paste0(projPath,"/", ExptName, "_metadataObject.RDS"))

```



## Make new bam-file for with only called cells
### R script to print filtered_cell_barcodes
```{r sinto_tsv}
dir.create(paste0(projPath,"/bam"))

cells<-
  # seur_hto@meta.data %>% rownames_to_column("barcode") %>%
  #  mutate(., is__cell_barcode = ifelse(
  #    HTO_classification.global =="Singlet",
  #    "1", "0")) %>%
  metadata %>%
  filter(is__cell_barcode=="1")%>% 
  dplyr::select(barcode, is__cell_barcode)

cells$is__cell_barcode<-paste0(ExptName, "_FilteredCells")

write_tsv(x = cells, file=paste0(projPath,"/bam/FilteredCellBarcodes.tsv"), col_names = FALSE)
```



## go to chapter 04.


```{r}
sessionInfo()

```
